---
title: "Tour through seqtime - Properties of time series generated with different ecological models"
author: "Karoline Faust"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{seqtime examples: Properties of time series from different models}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
  \usepackage[utf8]{inputenc}
---


```{r setup, include=FALSE}
# Global options
library(knitr)
opts_chunk$set(fig.path="figure_seqtime_tour/")
```

We start by loading the seqtime library.

```{r, message=FALSE, warning=FALSE}
library(seqtime)
```

Then, we generate a time series with the Ricker community model. For this, we have to specify the number of species N as well as the interaction matrix A that describes their interactions. We generate an interaction matrix with a specified connectance c at random. The connectance gives the percentage of realized arcs out of all possible arcs (arc means that the 
edge is directed).
The diagonal values, which represent intra-species competition, are set to a negative value. In addition, we introduce a high percentage of negative arcs to avoid explosions.

```{r}
N=50
A=generateA(N, c=0.1, d=-1)
A=modifyA(A,perc=70,strength="uniform",mode="negpercent")
# Generate a matrix using the algorithm by Klemm and Eguiluz to simulate a species network with a more realistic modular and scale-free structure. This takes a couple of minutes to complete.
#A=generateA(N, type="klemm", c=0.1)
```

Next, we generate uneven initial species abundances, summing to a total count of 1000.

```{r, fig.height = 6, fig.width = 6}
y=round(generateAbundances(N,mode=5))
names(y)=c(1:length(y))
barplot(y,main="Initial species abundances",xlab="Species",ylab="Abundance")
```

With the initial abundances and the interaction matrix, we can run a simulation of community dynamics with the Ricker model and plot the resulting time series. 

```{r, fig.height = 6, fig.width = 6}
# convert initial abundances in proportions (y/sum(y)) and run without a noise term (sigma=-1)
out.ricker=ricker(N,A=A,y=(y/sum(y)),K=rep(0.1,N), sigma=-1,tend=500)
tsplot(out.ricker,type="l",header="Ricker")
```

We can then analyse the community time series. First, we plot for each species
the mean against the variance. If a straight line fits well in log scale,
the Taylor law applies.

```{r, fig.height = 6, fig.width = 6}
# the pseudo-count allows to take the logarithm of species hat went extinct
ricker.taylor=seqtime::taylor(out.ricker, pseudo=0.0001, col="green", type="taylor")
```

Now we look at the noise types of the species simulated with the noise-free Ricker model. The only noise type identified is black noise.

```{r, fig.height = 6, fig.width = 6}
ricker.noise=identifyNoisetypes(out.ricker, abund.threshold = 0)
plotNoisetypes(ricker.noise)
```

Next, we run the SOI model on the same interaction matrix and initial abundances. For the example, we run it with only 500 individuals and 100 generations.

```{r, fig.height = 6, fig.width = 6}
out.soi=soi(N, I=500, A=A, m.vector=y, tend=100)
tsplot(out.soi,type="l",header="SOI")
```

The Taylor law fits far better to time series generated with the SOI than with the Ricker model according to the adjusted R2.

```{r, fig.height = 6, fig.width = 6}
soi.taylor=seqtime::taylor(out.soi, pseudo=0.0001, col="blue", type="taylor")
```

When we compute noise types in the community time series generated with SOI, we find that pink noise dominates.

```{r, fig.height = 6, fig.width = 6}
soi.noise=identifyNoisetypes(out.soi,abund.threshold=0)
plotNoisetypes(soi.noise)
```

Next, we generate a community time series with the Hubbell model, which describes neutral community dynamics. We set the number of species in the local and in the meta-community as well as the number of deaths to N and assign 1500 individuals. The immigration rate m is set to 0.1. We skip the first 500 steps of transient dynamics. 

```{r, fig.height = 6, fig.width = 6}
out.hubbell=simHubbell(N=N, M=N,I=1500,d=N, m.vector=(y/sum(y)), m=0.1, tskip=500, tend=1000)
tsplot(out.hubbell,type="l",header="Hubbell")
```

The neutral dynamics fits the Taylor law well:

```{r, fig.height = 6, fig.width = 6}
hubbell.taylor=seqtime::taylor(out.hubbell, pseudo=0.0001, col="blue", type="taylor")
```

The Hubbell time series is dominated by brown noise:

```{r, fig.height = 6, fig.width = 6}
hubbell.noise=identifyNoisetypes(out.hubbell,abund.threshold=0)
plotNoisetypes(hubbell.noise)
```

Finally, we generate a community with the Dirichlet Multinomial distribution, which in contrast to the three previous models does not introduce a dependency between time points.

```{r, fig.height = 6, fig.width = 6}
dm.uneven=simCountMat(N,samples=100,mode=5,k=0.05)
tsplot(dm.uneven,type="l",header="Dirichlet-Multinomial")
```

We plot its Taylor law.

```{r, fig.height = 6, fig.width = 6}
dm.uneven.taylor=seqtime::taylor(dm.uneven, pseudo=0.0001, col="orange", type="taylor", header="Dirichlet-Multinomial")
```

As expected, samples generated with the Dirichlet-Multinomial distribution do not display pink, brown or black noise. 

```{r, fig.height = 6, fig.width = 6}
dm.uneven.noise=identifyNoisetypes(dm.uneven,abund.threshold=0)
plotNoisetypes(dm.uneven.noise)
```

The evenness of the species proportion vector given to the Dirichlet-Multinomial distribution influences the slope of the Taylor law:

```{r, fig.height = 6, fig.width = 6}
dm.even=simCountMat(N,samples=100,mode=1)
dm.even.taylor=seqtime::taylor(dm.even, pseudo=0.0001, col="orange", type="taylor", header="Even Dirichlet-Multinomial")
```

## Noise simulation

For further examples on simulating noise, see [noise_simulations.md](noise_simulations.md).
