---
title: "Tour through seqtime - Noise types and Taylor laws in different data types"
author: "Karoline Faust"
date: "16 Dec 2015"
output: html_document
---

We start by loading the seqtime library.

```{r, message=FALSE, warning=FALSE}
library(seqtime)
```

Then, we generate a time series with Ricker's model. For this, we have to specify the number of species in the community as well as the interaction matrix that describes
their interactions. We generate an interaction matrix with a specified connectance at random.

```{r}
N=50
A=generateA(N, c=0.1, ignore.c=FALSE)
# Generate a matrix using the algorithm by Klemm and Eguiluz to simulate a species network with a realistic structure. This takes a couple of minutes to complete.
#A=generateA(N, type="klemm", c=0.1, ignore.c=FALSE)
```

The code below visualizes the species network encoded in A. Each node
has a self-arc, which represents intra-species competition.

```{r}
A.graph=graph.adjacency(A,mode="directed",weighted=TRUE)
plot(A.graph,layout=layout.grid(A.graph))
```

Now we generate uneven initial species abundances, summing to a total count of 1000.

```{r}
y=round(generateAbundances(N,mode=5))
names(y)=c(1:length(y))
barplot(y,main="Initial species abundances",xlab="Species",ylab="Abundance")
```

With the initial abundances and the interaction matrix, we can run the simulation with Ricker and plot the resulting time series. We set the carrying capacity to a very low value, to avoid explosions.

```{r}
out.ricker=ricker(N,A=A,y=y,K=rep(0.0005,N),tend=500)
tsplot(out.ricker,type="l",header="ricker")
```

We can then analyse the community time series. First, we plot for each species
the mean against the variance. If a straight line fits well in log scale,
the Taylor law applies. In case of the Ricker model, the Taylor law has a positive slope.

```{r}
# the pseudo-code allows to take the logarithm of species hat went extinct
taylor(out.ricker, pseudo=1, col="green", type="taylor")
```

Now we look at the noise types of the species simulated with the Ricker model. In this case, the only noise type that is clearly identified is brown noise.

```{r}
ricker.noise=identifyNoisetypes(out.ricker, abund.threshold = 0)
plot(ricker.noise)
```

Next, we run the SOC model on the same interaction matrix and initial abundances. For the example, we run it with only 500 individuals and 100 generations.

```{r}
out.soc=soc(N, I=500, A=A, m.vector=y, tend=100)
```

The Taylor law fits far better to species generated with SOC than with Ricker according to the R2.

```{r}
taylor(out.soc, pseudo=1, col="blue", type="taylor")
```

When we compute noise types in the community time series generated with SOC, we find that both pink and brown noise occurs.

```{r}
soc.noise=identifyNoisetypes(out.soc,abund.threshold=0)
plot(soc.noise)
```

Finally, we generate a community with the Dirichlet Multinomial distribution and plot
its Taylor law.

```{r}
dm.uneven=simCountMat(N,samples=100,mode=5,k=0.05)
taylor(dm.uneven, pseudo=1, col="orange", type="taylor", header="Dirichlet-Multinomial")
```

Only white noise occurs, as expected, since samples were generated independently with the Dirichlet-Multinomial distribution.

```{r}
# we relax the p-value of the linear fit in the log-scaled power spectrum
dm.uneven.noise=identifyNoisetypes(dm.uneven,pval.threshold = 1)
plot(dm.uneven.noise)
```

